---
title: "Data Cleanup"
---

Data Cleanup is being done in this notebook so it's easier to see results immediately. The Markdown portion of this isn't very important only the Notebook.

```{r}
library(tidyverse)
```


Initial cleanup removing unnecessary columns and rows
```{r}
library(readr)
School_full <- read_csv("Data/School_full.csv")

school_clean = School_full %>% 
  filter(membership_key == 188) %>% 
  select(lea_beds, nrc_code, subgroup_code, subgroup_name, enroll_cnt, grad_pct)
```

`school_clean` should now contain all the data from every school in NYS, with only the graduation data after students have had a full 6 years and 2 months to graduate, the longest a student can take to graduate and still be reported here.

```{r}
pred_male = 0.5
pred_white = 0.5
exc_disabled = 0.5
exc_ecd = 0.5
exc_esl = 0.5
exc_foster = 0.5
exc_homeless = 0.5
exc_migrant = 0.5
exc_armed = 0.5
```
These `pred` values represent how much of a school must have a particular trait for it to be considered to *predominantly* have that trait, in this case if a school is *predominantly* male is has more than 50% male students. The `exc` categories represent schools that have an *excess* of that particular trait, meaning that it falls above state or national averages (depending on which data was more readily available).


```{r}
num_schools = length(unique(school_clean$lea_beds))
num_schools

school_analysis = data.frame(matrix(nrow = num_schools, ncol = 13))

colnames(school_analysis) = c("beds_code", "enroll_count", "grad_pct", "nrc_code", "pred_male", "exc_disabled", "pred_white", "exc_ecd", "exc_esl", "exc_foster", "exc_homeless", "exc_migrant", "exc_armed")
```
This just creates the empty data frame that the school data will be transferred into.
There are 712 unique `beds` codes in this dataset, meaning 712 schools.

```{r}
n = 0 # this is a counter for which school we are on, it is names n just so it is easier to index
current_beds = 0

for (i in 1:length(school_clean$lea_beds)) {
  if(current_beds != school_clean$lea_beds[i]) {
    n = n + 1
    current_beds = school_clean$lea_beds[i]
    school_analysis$beds_code[n] = current_beds
  }
  
  group_code = school_clean$subgroup_code[i]
  switch(as.character(group_code),
         '1' = { # all students
           school_analysis$enroll_count[n] = school_clean$enroll_cnt[i]
           school_analysis$grad_pct[n] = school_clean$grad_pct[i]
           school_analysis$nrc_code[n] = school_clean$nrc_code[i]
           curr_schl_enrl = school_analysis$enroll_count[n]
         },
         '3' = { # male group
           school_analysis$pred_male[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > pred_male
         },
         '11' = { # students with disabilities
           school_analysis$exc_disabled[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > exc_disabled
         },
         '8' = { # white students
           school_analysis$pred_white[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > pred_white
         },
         '15' = { # economically disadvantaged
           school_analysis$exc_ecd[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > exc_ecd
         },
         '13' = { # English language learner
           school_analysis$exc_esl[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > exc_esl
         },
         '22' = { # in foster care
           school_analysis$exc_foster[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > exc_foster
         },
         '20' = { # homeless
           school_analysis$exc_homeless[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > exc_homeless
         },
         '17' = { # migrant
           school_analysis$exc_migrant[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > exc_migrant
         },
         '24' = { # parent in armed forces
           school_analysis$exc_armed[n] = school_clean$enroll_cnt[i]/curr_schl_enrl > exc_armed
         }
         )
}
```
This organizes all the data as we want, but leaves any category that a school had 0 students of unreported. We'll need to remove the schools that have `NA` for their grad percentage and then put in `FALSE` for the demographics that have an `NA`.


```{r}
school_analysis[1,2]

for (i in 1:712) {
  for (j in 5:13) {
    if(is.na(school_analysis[i,j])) school_analysis[i,j] = FALSE 
  }
}

school_analysis$grad_pct = as.numeric(substr(school_analysis$grad_pct, 1, 2))/100

school_analysis = school_analysis %>% 
  filter(!(is.na(grad_pct)))
```



